Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-10-06 paper/ifo-complexAudio.tex
combined notch and Wiener plots

We use the built-in MATLAB filter design toolbox in this work. 
The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}, respectively. We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
\end{center}
\caption{\label{fig:timeNotchWiener}
Timeseries results for the notch and Wiener filters. 
The first panel shows the original speech recording which was played from the speaker. 
The second panel shows the data from the optical microphone before any filters are applied. 
The third and fourth panels show results after applying the notch and Wiener filter respectively. 
Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
\end{center}
\caption{\label{fig:freqNotchWiener}
Frequency spectrum results for the notch and Wiener filters. 
The first and second panels show the results of applying the notch and Wiener filter respectively.
The third panel shows the spectrum after applying both the notch and Wiener filters. 
The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
}
\end{figure*}

@@

If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, if (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
\begin{equation}
\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
\end{equation}
where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
% What this sentence is trying to convey is not clear to me
%These assumptions are to be justified by the efficacy of the Wiener filter.
%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
%\begin{equation}
%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
%\end{equation}
%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
\label{eqn:wiener-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}\,,
\end{equation}
where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
%\begin{equation*}
%\left\{ 
\begin{eqnarray} 
r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
\end{eqnarray} 
%\right\}  
%\end{equation*}
%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.

@@

In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}. A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required. For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency. The effects of applying the Wiener filter to the time series and frequency spectrum are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener}, respectively. We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen visibly in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener}.
The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. The mains noise is all but removed and more voice information is retained. However, the recovered voice still sounds muffled and is not understandable.

>>>>

We use the built-in MATLAB filter design toolbox in this work. 
The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fourth row of Fig.~\ref{fig:NotchWienerCombined}.
We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=\textwidth]{figures/notch_and_wiener_superplot.pdf}
	\end{center}
	\caption{\label{fig:NotchWienerCombined}
	Time series (left column) and frequency spectrum (right column) results for the notch and Wiener filters.
	The first row shows the original speech recording.
	The second row shows the raw output from the optical microphone.
	The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
	The fourth, fifth, and sixth rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
	}
\end{figure*}

% \begin{figure*}
% \begin{center}
% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
% \end{center}
% \caption{\label{fig:timeNotchWiener}
% Timeseries results for the notch and Wiener filters. 
% The first panel shows the original speech recording which was played from the speaker. 
% The second panel shows the data from the optical microphone before any filters are applied. 
% The third and fourth panels show results after applying the notch and Wiener filter respectively. 
% Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
% }
% \end{figure*}

% \begin{figure*}
% \begin{center}
% \includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
% \end{center}
% \caption{\label{fig:freqNotchWiener}
% Frequency spectrum results for the notch and Wiener filters. 
% The first and second panels show the results of applying the notch and Wiener filter respectively.
% The third panel shows the spectrum after applying both the notch and Wiener filters. 
% The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
% }
% \end{figure*}

@@

If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, if (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
\begin{equation}
\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
\end{equation}
where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
% What this sentence is trying to convey is not clear to me
%% just trying to say that we can make the assumptions that we do because the Wiener filter appears to work well under them - J
%These assumptions are to be justified by the efficacy of the Wiener filter.
%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
%\begin{equation}
%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
%\end{equation}
%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
\label{eqn:wiener-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}\,,
\end{equation}
where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
%\begin{equation*}
%\left\{ 
\begin{eqnarray} 
r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
\end{eqnarray} 
%\right\}  
%\end{equation*}
%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.

@@

In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}. A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required. For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency. The time series and frequency spectrum after applying the Wiener filter are shown in the fifth row of Fig.~\ref{fig:NotchWienerCombined}.
We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:NotchWienerCombined}.
The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. The mains noise is all but removed and more voice information is retained. However, the recovered voice still sounds muffled and is not understandable.
